Effectiveness of community-based TURF-reserves in small-scale fisheries

### 2018-12-04 docs/paper/Villasenor-Derbez.Rmd
Comments in #19 have been adressed by explaining how TURF-reserves are similar to CBMRs, expanding on the rights. close #19

TURFs are a fisheries management tool in which a well defined group of fishers have exclusive access to an explicitly delimited portion of the ocean. They promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,mccay_2017}. On the other hand, no-take marine reserves (marine reserves from hereinafter) are areas where all extractive activities are off-limits. These can be implemented to protect biodiversity but also as fishery management tools that restrict fishing effort and gears and therefore aid in the recovery of marine stocks. Commonly known as TURF-reserves, the combination of these tools should in theory increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014,lester_2017}.

@@

TURF-reserve systems are inherently intrincate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact. This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009-hg}. There is a growing body of literature focusing on the relations between socioeconomic and governance structures and reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}. However, to our knowledge, no studies exist that evaluate TURF-reserves from both a social and ecological perspective.

@@

Before discussing our data collection methods and describing our analyses, our case studies warrant some more background. TURF-reserves can be created as community-based marine reserves, voluntarily established and enforced by local communities. This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{gelcich_2015,beger_2004,smallhornwest_2018}. Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}. However, marine reserves are difficult to enforce if they are not legally recognized, and fishers rely on the exclusive access granted by the TURF.

>>>>

TURFs are a fisheries management tool in which a well defined group of fishers have exclusive access to an explicitly delimited portion of the ocean. They promote a sense of stewardship and incentivise resource users to sustainably manage their resources \citep{gelcich_2008,mccay_2017}. On the other hand, no-take marine reserves (marine reserves from hereinafter) are areas where all extractive activities are off-limits. These can be implemented to protect biodiversity but also as fishery management tools that restrict fishing effort and gears and therefore aid in the recovery of marine stocks. Commonly known as TURF-reserves, the combination of these tools should in theory increase the benefits of spatial access rights allowing the maintenance of healthy resources \citep{afflerbach_2014,gelcich_2015,lester_2017}.

@@

TURF-reserve systems are inherently intrincate social-ecological systems, and their effectiveness must depend on how environmental and social factors combine and interact \citep{gelcich_2015}. This is especially important in social-ecological coastal systems dominated by close interactions and feedbacks between people and natural resources \citep{ostrom_2009-hg}. There is a growing body of literature focusing on the relations between socioeconomic and governance structures and reserve effectiveness \citep{halpern_2013,lpezangarita_2014,mascia_2017-m_,bergseth_2018}. However, to our knowledge, no studies exist that evaluate TURF-reserves from both a social and ecological perspective.

@@

Before discussing our data collection methods and describing our analyses, our case studies warrant some more background. Community-based marine reserves that are implemented within TURFs are a form of TURF-reserves, voluntarily established and enforced by local communities. This bottom-up approach increases compliance and self-enforcement, and reserves can yield benefits similar to systematically-designed reserves \citep{beger_2004,gelcich_2015,smallhornwest_2018}. Community-based spatial closures occur in different contexts, like the *kapu* or *ra’ui* areas in the Pacific Islands \citep{bohnsack_2004,johannes_2002}. However, marine reserves are difficult to enforce if they are not legally recognized, and fishers rely on the exclusive access granted by the TURF.

### 2018-12-04 docs/paper/Villasenor-Derbez.tex
Comments in #19 have been adressed by explaining how TURF-reserves are similar to CBMRs, expanding on the rights. close #19

\title[Community-based marine reserves]{Effectiveness of community-based TURF-reserves in small-scale fisheries}

@@

TURFs are a fisheries management tool in which a well defined group of
fishers have exclusive access to an explicitly delimited portion of the
ocean. TURFs promote a sense of stewardship and incentivise resource
users to sustainably manage their resources
\citep{gelcich_2008,mccay_2017}. On the other hand, no-take marine
reserves (marine reserves from hereinafter) are areas where all
extractive activities are off-limits. These can be implemented to
protect biodiversity but also as fishery management tools that restrict
fishing effort and gears and therefore aid in the recovery of marine
stocks. Commonly known as TURF-reserves, the combination of these tools
increase the benefits of spatial access rights allowing the maintenance
of healthy resources \citep{afflerbach_2014,lester_2017}.

Research on TURFs has shown that these areas have higher abundance of
targeted species than sites operating under open access
\citep{gelcich_2008,mccay_2014,mccay_2017}. Likewise, conservation
science has shown how marine reserves lead to increased biomass, species
richness, and abundance within the protected regions
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, and that these may
have a series of additional benefits like climate change mitigation,
protection from environmental variability, and fisheries benefits
\cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. The benefits
resulting from reserves established within TURFs are captured by the
group of fishers with exclusive access. Although in theory these systems
are successful \citep{costello_2010,smallhornwest_2018}, there is little
empirical evidence of their effectiveness and the drivers of their
success \citep{afflerbach_2014,lester_2017}.

@@

TURF-reserves can be created as community-based marine reserves,
voluntarily established and enforced by local communities. This
bottom-up approach increases compliance and self-enforcement, and
reserves can yield benefits similar to systematically-designed reserves
\citep{gelcich_2015-Gw,beger_2004-Y8,smallhornwest_2018}.
Community-based spatial closures occur in different contexts, like the
\emph{kapu} or \emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. However, marine reserves are
difficult to enforce if they are not legally recognized, and fishers
rely on the exclusive access granted by the TURF.

In an effort to bridge this normative gap, Mexican Civil Society
Organizations (CSOs) served as a link between fishers and government,
and created a legal framework that solves this governance issue
(\emph{i.e.} Fish Refuges; \citet{nom}). Fish refuges can be implemented
as temporal or partial reserves, which can protect one, some, or all
resources within their boundaries. One of the ways in which fishing
communities have takane advantage of this new tool is by implementing
marine reserves within their TURFs.

The common setup of community-based TURF-reserves in mexico is the
following. Fishers from a given community are assambled in fishing
cooperatives which have exclusive fishing rights over a spatially
delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
\ref{fig:map}A). Each TURF is exlusively fished by one cooperative, and
each community usually hosts no more than one cooperative. Fishing
cooperatives interested in implementing marine reserves work with CSOs
to implement marine reserves within their TURFs (\emph{i.e.}
TURF-reserves). Fishers then ask the government to grant legal
recognition to their TURF-reserves under the name of Fish refuges
following a series of studies outlined in \citet{nom}.

>>>>

\title[Community-based TURF-reserves]{Effectiveness of community-based TURF-reserves in small-scale fisheries}

@@

TURFs are a fisheries management tool in which a well defined group of
fishers have exclusive access to an explicitly delimited portion of the
ocean. They promote a sense of stewardship and incentivise resource
users to sustainably manage their resources
\citep{gelcich_2008,mccay_2017}. On the other hand, no-take marine
reserves (marine reserves from hereinafter) are areas where all
extractive activities are off-limits. These can be implemented to
protect biodiversity but also as fishery management tools that restrict
fishing effort and gears and therefore aid in the recovery of marine
stocks. Commonly known as TURF-reserves, the combination of these tools
should in theory increase the benefits of spatial access rights allowing
the maintenance of healthy resources
\citep{afflerbach_2014,lester_2017}.

Research on TURFs has shown that these areas have higher abundance of
targeted species than sites operating under open access
\citep{gelcich_2008,mccay_2014,mccay_2017}. Likewise, conservation
science has shown how marine reserves lead to increased biomass, species
richness, and abundance within the protected regions
\citep{lester_2009-Ks,giakoumi_2017-V2,sala_2017-69}, and that these may
have a series of additional benefits like climate change mitigation,
protection from environmental variability, and fisheries benefits
\cite{roberts_2017-J9,micheli_2012-EU,krueck_2017-J1}. The benefits
resulting from reserves established within TURFs should be captured
exclusively by the group of fishers with exclusive access. Although in
theory these systems are successful
\citep{costello_2010,smallhornwest_2018}, there is little empirical
evidence of their effectiveness and the drivers of their success
\citep{afflerbach_2014,lester_2017}.

@@

Before discussing our data collection methods and describing our
analyses, our case studies warrant some more background. TURF-reserves
can be created as community-based marine reserves, voluntarily
established and enforced by local communities. This bottom-up approach
increases compliance and self-enforcement, and reserves can yield
benefits similar to systematically-designed reserves
\citep{gelcich_2015,beger_2004,smallhornwest_2018}. Community-based
spatial closures occur in different contexts, like the \emph{kapu} or
\emph{ra'ui} areas in the Pacific Islands
\citep{bohnsack_2004,johannes_2002}. However, marine reserves are
difficult to enforce if they are not legally recognized, and fishers
rely on the exclusive access granted by the TURF.

In an effort to bridge this normative gap, Mexican Civil Society
Organizations (CSOs) served as a link between fishers and government,
and created a legal framework that solves this governance issue
(\emph{i.e.} Fish Refuges; \citet{nom}). Fish refuges can be implemented
as temporal or partial reserves, which can protect one, some, or all
resources within their boundaries. One of the ways in which fishing
communities have takane advantage of this new tool is by implementing
marine reserves within their TURFs. Our work focuses on some of these
community-based TURF-reserves that occur in small-scale fisheries.

The common setup of community-based TURF-reserves in Mexico is the
following. Fishers from a given community are assambled in fishing
cooperatives which have exclusive fishing rights over a spatially
delimited area (\emph{i.e.} TURFs shown as blue polygons in Fig
\ref{fig:map}A). Each TURF is exlusively fished by one cooperative, and
each community usually hosts no more than one cooperative. Fishing
cooperatives interested in implementing marine reserves work with CSOs
to implement marine reserves within their TURFs (\emph{i.e.}
TURF-reserves). Fishers then ask the government to grant legal
recognition to their TURF-reserves under the name of Fish refuges
following a series of studies outlined in \citet{nom}.
